### 2017-10-06 paper/appendix.tex
appendix and figures

\subsection{DNN Training and Testing}
We have trained and tested the deep neural network with several
different track conditions, different combinations of input
data, and different hyper parameters. In the following, we describe
details on two of the training methods that performed reasonably well.

In the first method, we trained the neural network model across a set 
of 30 completed runs on the track seen in Figure~\ref{fig:track} by a
human pilot. Half of the runs saw the car driving one way along the
track, while the remaining half were of the car driving in the
opposite direction on the track.
In total, we collected 2,556 frames for training and 2,609 
frames for validation.
The weights of the network is initialized using Xavier
initializer~\cite{Glorot2010}, which is known to provide better
initial values than the random weight assignment method.
In each training step, we use a batch
size of 100 frames, which are randomly selected among all the
collected training images, to optimize the network.
We repeat this across 2,000 training steps. When a model was trained
with the  aformentioned data, the training loss was 0.0188 and the
validation  loss was 0.0132.
The change of the loss value over the course of model training can be
seen in Figure~\ref{fig:modelloss}.

\begin{figure*}[t]
  \centering
  \includegraphics[width=.9\textwidth]{figs/TrainingLoss}
  \caption{Change in loss value throughout training.}
  \label{fig:modelloss}
\end{figure*}

In the second method, we use the same data and parameters above except
that now images are labeled as 'curved' and 'straight' and we pick an
equal number of images from each category at each training step to
update the model. In other words, we try to remove bias in selecting
images. We find that the car performed better in practice by applying
this approach as the car displayed a greater ability to stay in the
center of the track (on the white tape).
However, we find that there is huge discrepency between the training
loss and the validation loss as the former was 0.09 while the latter
was 0.0869---a 10X difference---indicating that the model suffers from
the overfitting problem.

We continue to investigate ways to achieve better prediction accuracy
in training the network as well as improving the performance of the RC
car platform, especially on precise steering angle control.

%% We believe, though, that this can be
%% rectified through the collection of additional data for training and
%% validation.

%% \begin{figure*}[h]
%%   \centering
%%   \includegraphics[width=.9\textwidth]{figs/graph_run=.pdf}
%%   \caption{ Network visualization. }
%%   \label{fig:}
%% \end{figure*}
%% \fixme{Is all of the network graph necessary or only certain parts?}

\subsection{System-level Factors Affecting Real-Time Performance}
In using the Raspberry Pi 3 platform, there are a
few system-level factors, namely power supply and temperature, that
need to be considered to achieve consistent and high real-time
performance.

%% Specifically, these issues all have the potential to negatively
%% affect the cpu clock speed/frequency, which would result in decreased
%% performance. 
In all our experiments on the Raspberry Pi 3, the CPU  operated at a
preferred clock speed of 1.2 GHz. However, without care, it is
possible for the CPU may be operating at a lower frequency.

An important factor is CPU thermal throttling, which can affect CPU
clock speed if the CPU temperature is too high (Pi 3's firmware is
configured to throttle at 85C).
DNN model operations are computationally intensive, thus it is
possible for the temperature of the CPU to rise quickly. This can be
especially problematic in situations where multiple DNN models are
running simultaneously on the Pi 3.
If the temperature reaches the threshold, the Pi 3's thermal
throttling kicks in and decreases the clock speed down to 600MHz---a
half of the maximum 1.2GHz---so that the CPU's temperature stays at a
safe level.
We found that without proper cooling solutions (heatsink or fan), 
prolonged use of the system would result in CPU frequency decrease
that may affect evaluation.

Another factor to consider is power supply. From our experiences, the
Pi 3 frequency throttling also kicks in when the power source can not
provide the required minimum of 2A current.
In experiments conducted with a power supply that only provided 1 Amp,
the Pi was unable to sustain a 1.2 GHz clock speed and, instead,
fluctuated between operating at 600 MHz and 1.2 GHz. As a result, it
is necessary, or at least highly recommended, that the power supply
used for the Raspberry Pi 3 be capable of outputting 2 Amps, otherwise
optimal performance isn't guaranteed.

Our initial experiment results suffered from these issues after which
we always carefully monitored the current operating frequencies of the
CPU cores during the experiments to ensure the correctness and
repeatability of the results.

%% Thus the
%% Raspberry Pi may not be suited for prolonged use, especially in cases
%% where the workload is relatively larger, such as running multiple
%% models. Rather, the Pi seems to be better suited for running in set
%% periods, after which it is turned  off or made idle so that the cpu is
%% allowed time to cool down.