### 2018-04-15 paper/main.tex

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{adjustbox}
\usepackage{subcaption}
\usepackage{listings}
\usepackage{color}
\usepackage{hyperref}
\usepackage{fixme}
%\usepackage{authblk}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{DeepPicar: A Low-cost Deep Neural Network-based Autonomous Car}

\author{\IEEEauthorblockN{Michael G. Bechtel}
\IEEEauthorblockA{\textit{Electrical Engineering and Computer Science} \\
\textit{University of Kansas}\\
Lawrence, USA \\
mbechtel@ku.edu}
\and
\IEEEauthorblockN{Elise McEllhiney}
\IEEEauthorblockA{\textit{Electrical Engineering and Computer Science} \\
\textit{University of Kansas}\\
Lawrence, USA \\
e908m428@ku.edu}
\and
\IEEEauthorblockN{Minje Kim}
\IEEEauthorblockA{\textit{Electrical Engineering and Computer Science} \\
\textit{Indiana University Bloomington}\\
Bloomington, USA \\
minje@indiana.edu}
\and
\IEEEauthorblockN{Heechul Yun}
\IEEEauthorblockA{\textit{Electrical Engineering and Computer Science} \\
\textit{University of Kansas}\\
Lawrence, USA \\
heechul.yun@ku.edu}
}

\maketitle

\begin{abstract}
We present DeepPicar, a low-cost deep neural network (DNN) based
autonomous car platform. DeepPicar is a small scale
replication of a real self-driving car called Dave-2 by NVIDIA, which
drove on public roads using a deep convolutional neural network (CNN), 
that takes images from a front-facing camera as input and produces
car steering angles as output. DeepPicar uses almost the exact same 
network architecture---9 layers, 27 million connections and 250K
parameters---and can be trained to drive itself, in real-time, using a
web camera and a modest Raspberry Pi 3 quad-core platform.
Using DeepPicar, we analyze the Pi 3's computing capabilities to 
support end-to-end deep learning based real-time control of autonomous
vehicles. We also systematically compare other contemporary embedded
computing platforms using the DeepPicar's CNN based real-time control
software as a workload. 
We find all tested platforms, including the Pi 3, are capable of
supporting deep-learning based real-time control, from 20 Hz up to 100
Hz depending on hardware platform. 
However, shared resource contention remains an
important issue that must be considered in applying deep-learning
models on shared memory based embedded computing platforms.
\end{abstract}

\begin{IEEEkeywords}
Real-time, Autonomous car, Convolutional neural network, Case study
\end{IEEEkeywords}

%-------------------------------------------------------------------------
\input{intro}
\input{background}
\input{overview}
\input{evaluation}
%\input{evaluation-cachepart}
\input{comparison}
%\input{discussion}
\input{related}
\input{conclusion}
%-------------------------------------------------------------------------
\bibliographystyle{abbrv}
\bibliography{reference}

\appendix
\input{appendix}
\end{document}