### 2018-04-16 paper/intro.tex

\section{Introduction} \label{sec:intro}

% broad context:
% - advance in ai sparked interests in the robotics application, such as
%   self-driving cars.
% - in particular, deep neural network models are increasingly used
%   for perception and control of a vehicle. say. AI workloads.
%
%
Autonomous cars have been a topic of increasing interest in recent
years as many companies are actively developing related hardware
and software technologies toward fully autonomous driving capability with
no human intervention. Deep neural networks (DNNs) have been
successfully applied in various perception and control tasks in
recent years.  They are important workloads for autonomous vehicles
as well. For example, Tesla Model S was known to use a specialized
chip (MobileEye EyeQ), which used a deep neural network for vision-based
real-time obstacle detection and avoidance. More recently, researchers
are investigating DNN based end-to-end control of
cars~\cite{Bojarski2016} and other robots. It is expected that more
DNN based Artificial Intelligence workloads may be used in future
autonomous vehicles.

% big problem
Executing these AI workloads on an embedded computing platform 
poses several additional challenges. First, many AI workloads in vehicles 
are computationally demanding and have strict real-time requirements. 
For example, latency in a vision-based object
detection task may be directly linked to the safety of the vehicle. This
requires a high computing capacity as well as the means to guaranteeing
the timings. On the other hand, the computing hardware platform must
also satisfy cost, size, weight, and power constraints, which require a
highly efficient computing platform. These two conflicting
requirements complicate the platform selection process as observed in
~\cite{Otterness2017}.

%% For example, while today's self-driving car
%% prototype equip more \$100,000
%% of computers and sensors~\cite{juliussen2014emerging}, a study
%% found that aveage consumers are willing to pay much less amount of
%% extra cost for a self-driving capability~\cite{Daziano2017}.
%% https://qz.com/924212/what-it-really-costs-to-turn-a-car-into-a-self-driving-vehicle/

% related work and remaining problems
To understand what kind of computing hardware is needed for AI
workloads, we need a testbed and realistic workloads. While using a
real car-based testbed would be most ideal, it is not only highly
expensive, but also poses serious safety concerns that hinder
development and exploration. Therefore, there is a need for safer and
less costly testbeds.

%% There are already several relatively inexpensive RC-car
%% based testbeds, such as MIT's 
%% RaceCar~\cite{shin2017project} and UPenn's F1$/$10 racecar~\cite{upennf1tenth}.
%% However, these RC-car testbeds still cost more than \$3,000, requiring
%% considerable investment.

%% % our goals
%% Instead, we want to build a low cost testbed that still employs the
%% state-of-the art AI technologies. Specifically, we focus on an end-to-end
%% deep learning based real-time control system,
%% which was developed for a real self-driving car, NVIDIA
%% DAVE-2~\cite{Bojarski2016}, and uses the same methodology on a
%% smaller and less costly setup. In developing the testbed, our
%% goals are (1) to analyze real-time issues in DNN based end-to-end
%% control; and (2) to evaluate real-time performance of contemporary embedded
%% platforms for such a workload.

% DeepPicar introduction
In this paper, we present DeepPicar, a low-cost autonomous car
testbed for research. From a hardware perspective,
DeepPicar is comprised of a Raspberry Pi 3 Model B quad-core
computer, a web camera and a RC car, all of which are affordable
components (less than \$100 in total).
The DeepPicar, however, employs a state-of-the-art AI
technology, namely deep learning based end-to-end real-time control,
which utilizes a deep convolutional neural network (CNN).
The CNN receives an image frame from a single forward
looking camera as input and generates a predicted steering angle
value as output at each control period in \emph{real-time}.
The CNN has 9 layers, about 27 million connections
and 250 thousand parameters (weights).
DeepPicar's CNN architecture is identical to that of NVIDIA's
real-sized self-driving car, called DAVE-2~\cite{Bojarski2016}, which
drove on public roads without human driver's intervention while only
using the CNN, which runs on a significantly more powerful computing
platform (Drive PX~\cite{drivepx}).

%% , which uses a significantly more
%% powerful computer (Drive PX computer~\cite{drivepx}) than a Raspberry
%% Pi~3.
%% We chose to use a Pi 3 not only because it is affordable, but 
%% also because it is representative of today's mainstream low-end
%% embedded multicore platforms found in smartphones and other embedded
%% devices.

%% Other than the difference in scale (RC car vs. real car), the only other
%% differences between the two systems---from the computing
%% perspective---are that our system is implemented in
%% TensorFlow~\cite{abadi2016tensorflow} and runs on a Raspberry Pi 3
%% whereas NVIDIA's DAVE-2 systems is implemented in Torch
%% 7~\cite{collobert2011torch7} and runs on a Drive PX computer (NVIDIA's
%% automotive specialized computing system~\cite{drivepx}), which is more
%% powerful but also more expensive.

% how we trained (maybe moved to a later section)
%% We apply a standard imitation learning methodology to train the car to
%% follow tracks on the ground. We collect data for
%% training and validation by manually
%% controlling the RC car and recording the vision (from the webcam
%% mounted on the RC-car) and the human control inputs. We then train the
%% network offline using the collected data on a desktop computer, which
%% is equipped with a NVIDIA GTX 1050 Ti GPU. Finally, the trained network is copied
%% back to the Raspberry Pi 3, which is then used to perform inference
%% operations---locally on the Pi 3---in the car's main control loop in
%% real-time. For real-time control, each inference operation must
%% be completed within the desired control period (e.g., 33.$\overline{\mbox{33}}$ ms
%%  period for 30 Hz control frequency).
% how we evaluated (in terms real-time performance)

% what are our findings?
Using DeepPicar, we systematically analyze its real-time
capabilities in the context of deep-learning based real-time
control, especially on real-time deep neural network \emph{inferencing}.
We also evaluated other, more powerful, embedded computing
platforms to better understand achievable real-time performance of
DeepPicar's deep-learning based control system and the impact of
computing hardware architectures.

% Findings:
%
% - we find DNN processing is an ideal real-time workload from the
%   perspective of timing predictability as the amount of computation
%   needed is fixed at design time and doesn't change over input.
%
% - real-time processing of non-trivial DNN inferencing is feasible
%   on today's embedded multicore platform, even as inexpensive as
%   raspberry pi3.
%
% - all tested platforms have enough computing capacity to consolidate 
%   multiple DeepPicar's CNN workloads or other compute intensive tasks
%   simultaneously.
%
% - however, when consolidating multiple workloads shared resource
%   contention is an important issue. we carefully analyzed
%   sensitivity of DeepPicar's workload with respect to shared
%   hardware resources.
%
% - in particular, we find DNN processing is highly sensitive to
%   memory performance. performance severely affected when memory
%   bandwidth is contented.
%
% - however, we find that dnn processing is not sensitive to cache
%   space. we find applying cache partitioning does not help improve
%   isolation of DNN workload when memory bandwidth is contended.
From the systematic study, we have made a number of interesting
observations.
% predictability
First, we find that CNN inferencing is highly time predictable as the
amount of computation needed to complete a 
single inference is fixed at the network architecture design time and
does not change at runtime over different inputs (e.g., different image
frames). This predictable timing behavior is obviously a desirable
property for real-time systems.
%% , which gives another reason for the
%% real-time community to be interested in the use of DNNs.

% performance on low-cost embedded multicore
Second, we find that real-time processing of CNN inferencing is
feasible on today's embedded computing platforms, even one   
as inexpensive as the Raspberry Pi 3. For example, just one Cortex-A53
core of the Raspberry Pi3 is able to complete the CNN inferencing task
in less than 50ms, sufficient to support 20 Hz control loop. 
%% We find that the control loop 
%% completes in under 33.$\overline{\mbox{33}}$ ms, or 30 Hz, using just two 
%% cores of the Raspberry Pi 3's quad-core CPU (w/o using its GPU), and can
%% do so 100\% of the time; or we can achieve 20 Hz control frequency using
%% just one core.
%% Other tested embedded platforms, the Intel UP and NVIDIA TX2, offer even
%% better performance, and are capable of supporting deep-learning based 
%% real-time control over 100 Hz control frequency on the TX2 when its
%% embedded GPU is used.
We find that all the tested embedded computing platforms have 
enough computing capacity to consolidate multiple instances of
DeepPicar's DNN workload.

% all platforms have enough capacity to consolidate
Third, contention in shared hardware resources remains an important
issue when consolidating multiple workloads. 
We observe up to 11.6X control
loop execution time increase due to shared resource contention when
memory intensive  applications are co-scheduled on idle CPU cores.
We find that DeepPicar's inferencing task is highly
sensitive to memory bandwidth, but not cache space.

%% ; we show that applying cache partitioning has negligible impact
%% in providing timing isolation on consolidated scenarios.
%% This finding
%% suggests that in order to guarantee real-time performance of DNN-based
%% control applications, solutions to guarantee their memory performance
%% are more important.

% contributions
The {\bf contributions} of this paper are as follows:
\begin{itemize}
  \item We present DeepPicar, a low-cost autonomous car testbed using
    a state-of-the-art CNN for end-to-end real-time control, to lower
    economic and safety-related barriers to the study autonomous
    cars.~\footnote{The source code, dataset, and build instructions
      of DeepPicar can be found at:\fixme{Michael github url}}

  \item We provide extensive systematic experimental results using
    DeepPicar's CNN inferencing workload on a number of representative
    contemporary embedded computing platforms. The results provide
    insights for real-time processing of the CNN workload and the
    effect of consolidation.

    %% While larger embedded
    %% platforms have been evaluated, as in  \cite{Otterness2017}, to the
    %% best of our knowledge, no prior work investigated the
    %% impacts of shared resource contention on DNN inferencing workloads
    %% in the context of smaller embedded platforms like the Raspberry Pi 3.
\end{itemize}

The remaining sections of the paper are as follows.
Section~\ref{sec:background}
provides a background of applications in autonomous driving and related works.
Section~\ref{sec:overview} gives an overview of the DeepPicar
platform, including the high-level system and the methods used for
training and inference. Section~\ref{sec:evaluation} presents our
evaluation of the platform and how different factors can affect
performance. Section~\ref{sec:comparison} offers a comparison between  
the Raspberry Pi 3 and other embedded computing platforms to 
determine their suitably for autonomous driving research.
%% We discuss more detailed training methods and system-level issues
%% in~\ref{sec:discussion}. 
We review related work in
Section~\ref{sec:related} and conclude in
Section~\ref{sec:conclusion}.