### 2018-04-20 paper/intro.tex
intro revision

\section{Introduction} \label{sec:intro}

% broad context:
% - advance in ai sparked interests in the robotics application, such as
%   self-driving cars.
% - in particular, deep neural network models are increasingly used
%   for perception and control of a vehicle. say. AI workloads.
%
%
Autonomous cars have been a topic of increasing interest in recent
years as many companies are actively developing related hardware
and software technologies toward fully autonomous driving capability with
no human intervention. Deep neural networks (DNNs) have been
successfully applied in various perception and control tasks in
recent years.  They are important workloads for autonomous vehicles
as well. For example, Tesla Model S was known to use a specialized
chip (MobileEye EyeQ), which used a deep neural network for vision-based
real-time obstacle detection and avoidance. More recently, researchers
are investigating DNN based end-to-end control of
cars~\cite{Bojarski2016} and other robots. It is expected that more
DNN based artificial intelligence (AI) workloads may be used in future
autonomous vehicles.

% big problem
Executing these AI workloads on an embedded computing platform 
poses several additional challenges. First, many AI workloads in vehicles 
are computationally demanding and have strict real-time requirements. 
For example, latency in a vision-based object
detection task may be directly linked to the safety of the vehicle. This
requires a high computing capacity as well as the means to guaranteeing
the timings. On the other hand, the computing hardware platform must
also satisfy cost, size, weight, and power constraints, which require a
highly efficient computing platform. These two conflicting
requirements complicate the platform selection process as observed in
~\cite{Otterness2017}.

%% For example, while today's self-driving car
%% prototype equip more \$100,000
%% of computers and sensors~\cite{juliussen2014emerging}, a study
%% found that aveage consumers are willing to pay much less amount of
%% extra cost for a self-driving capability~\cite{Daziano2017}.
%% https://qz.com/924212/what-it-really-costs-to-turn-a-car-into-a-self-driving-vehicle/

% related work and remaining problems
To understand what kind of computing hardware is needed for AI
workloads, we need a testbed and realistic workloads. While using a
real car-based testbed would be most ideal, it is not only highly
expensive, but also poses serious safety concerns that hinder
development and exploration. Therefore, there is a need for safer and
less costly testbeds.

%% There are already several relatively inexpensive RC-car
%% based testbeds, such as MIT's 
%% RaceCar~\cite{shin2017project} and UPenn's F1$/$10 racecar~\cite{upennf1tenth}.
%% However, these RC-car testbeds still cost more than \$3,000, requiring
%% considerable investment.

%% % our goals
%% Instead, we want to build a low cost testbed that still employs the
%% state-of-the art AI technologies. Specifically, we focus on an end-to-end
%% deep learning based real-time control system,
%% which was developed for a real self-driving car, NVIDIA
%% DAVE-2~\cite{Bojarski2016}, and uses the same methodology on a
%% smaller and less costly setup. In developing the testbed, our
%% goals are (1) to analyze real-time issues in DNN based end-to-end
%% control; and (2) to evaluate real-time performance of contemporary embedded
%% platforms for such a workload.

% DeepPicar introduction
In this paper, we present DeepPicar, a low-cost autonomous car
testbed for research. From a hardware perspective,
DeepPicar is comprised of a Raspberry Pi 3 Model B quad-core
computer, a web camera and a RC car, all of which are affordable
components (less than \$100 in total).
The DeepPicar, however, employs a state-of-the-art AI
technology, namely end-to-end deep learning based real-time control,
which utilizes a deep convolutional neural network (CNN).
The CNN receives an image frame from a single forward
looking camera as input and generates a predicted steering angle
value as output at each control period in \emph{real-time}.
The CNN has 9 layers, about 27 million connections
and 250 thousand parameters (weights).
DeepPicar's CNN architecture is identical to that of NVIDIA's
real-sized self-driving car, called DAVE-2~\cite{Bojarski2016}, which
drove on public roads without human driver's intervention while only
using the CNN.
%% Note that, DAVE-2 run on a significantly more powerful computing
%% platform (Drive PX~\cite{drivepx}).

%% , which uses a significantly more
%% powerful computer (Drive PX computer~\cite{drivepx}) than a Raspberry
%% Pi~3.
%% We chose to use a Pi 3 not only because it is affordable, but 
%% also because it is representative of today's mainstream low-end
%% embedded multicore platforms found in smartphones and other embedded
%% devices.

%% Other than the difference in scale (RC car vs. real car), the only other
%% differences between the two systems---from the computing
%% perspective---are that our system is implemented in
%% TensorFlow~\cite{abadi2016tensorflow} and runs on a Raspberry Pi 3
%% whereas NVIDIA's DAVE-2 systems is implemented in Torch
%% 7~\cite{collobert2011torch7} and runs on a Drive PX computer (NVIDIA's
%% automotive specialized computing system~\cite{drivepx}), which is more
%% powerful but also more expensive.

% how we trained (maybe moved to a later section)
%% We apply a standard imitation learning methodology to train the car to
%% follow tracks on the ground. We collect data for
%% training and validation by manually
%% controlling the RC car and recording the vision (from the webcam
%% mounted on the RC-car) and the human control inputs. We then train the
%% network offline using the collected data on a desktop computer, which
%% is equipped with a NVIDIA GTX 1050 Ti GPU. Finally, the trained network is copied
%% back to the Raspberry Pi 3, which is then used to perform inference
%% operations---locally on the Pi 3---in the car's main control loop in
%% real-time. For real-time control, each inference operation must
%% be completed within the desired control period (e.g., 33.$\overline{\mbox{33}}$ ms
%%  period for 30 Hz control frequency).
% how we evaluated (in terms real-time performance)

% what are our findings?
Using DeepPicar, we systematically analyze its real-time
capabilities in the context of end-to-end deep-learning based real-time
control, especially on real-time \emph{inferencing} of the CNN.
We also evaluated other, more powerful, embedded computing
platforms to better understand achievable real-time performance of
DeepPicar's deep-learning based control system and the impact of
computing hardware architectures.
From the systematic study, we want to answer the following questions:
(1) How well does the CNN based real-time inferencing task perform on
contemporary embedded multicore computing platforms? (2) How
susceptible is the CNN inferencing task to contention in shared
hardware resources (e.g., cache and DRAM) when multiple tasks/models
are consolidated? (3) Are the existing state-of-the-art shared resource
isolation techniques effective in protecting real-time performance of
the CNN inferencing task?

% Findings:
%
% - we find DNN processing is an ideal real-time workload from the
%   perspective of timing predictability as the amount of computation
%   needed is fixed at design time and doesn't change over input.
%
% - real-time processing of non-trivial DNN inferencing is feasible
%   on today's embedded multicore platform, even as inexpensive as
%   raspberry pi3.
%
% - all tested platforms have enough computing capacity to consolidate 
%   multiple DeepPicar's CNN workloads or other compute intensive tasks
%   simultaneously.
%
% - however, when consolidating multiple workloads shared resource
%   contention is an important issue. we carefully analyzed
%   sensitivity of DeepPicar's workload with respect to shared
%   hardware resources.
%
% - in particular, we find DNN processing is highly sensitive to
%   memory performance. performance severely affected when memory
%   bandwidth is contented.
%
% - however, we find that dnn processing is not sensitive to cache
%   space. we find applying cache partitioning does not help improve
%   isolation of DNN workload when memory bandwidth is contended.
Our observations are as follows.
First, we find that real-time processing of the CNN inferencing is
feasible on contemporary embedded computing platforms, even one   
as inexpensive as the Raspberry Pi 3. Second, while consolidating
multiple CNN models and task are feasible but the impact of shared
resource contention can be catastrophic in the worst-case as we
observe up to 11.6X slowdown even when the CNN model was given a
dedicated core. Third, using the state-of-the-art OS-level cache
partitioning and memory bandwidth control
mechanisms~\cite{yun2014rtas,Yun2013}, we find that cache space has a
minimal performance impact to the CNN workload while memory
performance has a significant impact.

%% % predictability
%% First, we find that CNN inferencing is highly time predictable as the
%% amount of computation needed to complete a 
%% single inference is fixed at the network architecture design time and
%% does not change at runtime over different inputs (e.g., different image
%% frames). This predictable timing behavior is obviously a desirable
%% property for real-time systems.
%% %% , which gives another reason for the
%% %% real-time community to be interested in the use of DNNs.
% performance on low-cost embedded multicore
%% For example, just one Cortex-A53
%% core of the Raspberry Pi3 is able to complete the CNN inferencing task
%% in less than 50ms, sufficient to support 20 Hz control loop. 
%% We find that the control loop 
%% completes in under 33.$\overline{\mbox{33}}$ ms, or 30 Hz, using just two 
%% cores of the Raspberry Pi 3's quad-core CPU (w/o using its GPU), and can
%% do so 100\% of the time; or we can achieve 20 Hz control frequency using
%% just one core.
%% Other tested embedded platforms, the Intel UP and NVIDIA TX2, offer even
%% better performance, and are capable of supporting deep-learning based 
%% real-time control over 100 Hz control frequency on the TX2 when its
%% embedded GPU is used.
%% We find that all the tested embedded computing platforms have 
%% enough computing capacity to consolidate multiple instances of
%% DeepPicar's DNN workload.

% all platforms have enough capacity to consolidate
%% Third, contention in shared hardware resources remains an important
%% issue when consolidating multiple workloads. 
%% We observe up to 11.6X control
%% loop execution time increase due to shared resource contention when
%% memory intensive  applications are co-scheduled on idle CPU cores.
%% We find that DeepPicar's inferencing task is highly
%% sensitive to memory bandwidth, but not cache space.

%% ; we show that applying cache partitioning has negligible impact
%% in providing timing isolation on consolidated scenarios.
%% This finding
%% suggests that in order to guarantee real-time performance of DNN-based
%% control applications, solutions to guarantee their memory performance
%% are more important.

% contributions
This paper makes the following {\bf contributions}:
\begin{itemize}
  \item We present DeepPicar, a low-cost autonomous car testbed, which
    employs a state-of-the-art CNN based end-to-end
    real-time control~\footnote{The source code, datasets, and build
      instructions of DeepPicar can be found at:
      \url{https://github.com/mbechtel2/DeepPicar-v2}}.
  \item We provide extensive empirical performance evaluation results
    of the CNN inferencing workload on a number of contemporary
    embedded computing platforms.
  \item We apply the state-of-the-art shared resource isolation
    techniques and evaluate their effectiveness in protecting the
    inferencing workload in consolidated settings.
\end{itemize}

The remainder of the paper is organized as follows.
Section~\ref{sec:background} provides a background on the application
of nueral networks in autonomous driving.
Section~\ref{sec:overview} gives an overview of the DeepPicar
testbed. Section~\ref{sec:evaluation} presents extensive real-time
performance evaluation results we collected on the testbed.
Section~\ref{sec:comparison} offers a comparison between  
the Raspberry Pi 3 and other embedded computing platforms.
%% We discuss more detailed training methods and system-level issues
%% in~\ref{sec:discussion}. 
We review related work in
Section~\ref{sec:related} and conclude in
Section~\ref{sec:conclusion}.